### 2011-11-19 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: merging Lydka round 1.

\begin{define}[Node equality]
	$v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
\end{define}

\begin{define}[Value equality]
	$v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
\end{define}

@@

%\section{XML Keys} - is this necessary at all?

%\subsection{According to Specification}

% - compare them to ID attributes

%\subsection{According to \cite{keX}}

% - show how a XML key is defined using the keX notation % this is a terrible sentence

@@

\begin{define}[Attribute mapping]
	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $\mathcal{I} \times \mathcal{I}$ relation defined by
	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}\]
\end{define}

@@

\begin{define}[Image of the attribute mapping]
	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$
\end{define}

@@

\begin{define}[Candidate attribute mapping]
An attribute mapping $m$ is a \textit{candidate attribute mapping} if it is an injective function, that is,
\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|\]
\end{define}

@@

\begin{define}[ID set]
A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset\]
\end{define}

@@

Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}.
\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.

@@

This definition of weight for AMs or AM sets comes from \cite{fidax} again.

Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.

@@

\begin{define}[Support]
\textit{Support} of an attribute mapping $m$ is defined as follows.
\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}\]
\end{define}

@@

\begin{define}[Coverage]
\textit{Coverage} of an attribute mapping $m$ is defined as follows.
\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|\]
\end{define}

@@

\begin{define}[Weight]
For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows.
\[weight(m) = \alpha . \phi(m) + \beta . \chi(m)\]
For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs.
\[weight(S) = \sum_{m \in S} weight(m)\]
\end{define}

@@

\begin{define}[Independent set]
Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff the following holds.
\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E \]
\end{define}

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v)\]
\end{define}

@@

It can be written in the following form (minimization version is possible too).

\begin{eqnarray*}
\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
\mathbf{x} & \geqslant & 0 \\
\end{eqnarray*}
\[
\mathbf{x} =
\begin{bmatrix}
x_1 \\
x_2 \\
\ldots \\
x_n \\
\end{bmatrix},
\mathbf{c} = 
\begin{bmatrix}
c_1 \\
c_2 \\
\ldots \\
c_n \\
\end{bmatrix},
\mathbf{b} =
\begin{bmatrix}
b_1 \\
b_2 \\
\ldots \\
b_m \\
\end{bmatrix}
\]

@@

Another way to write this formulation is in the following.

@@

\begin{define}[Mixed integer problem]
	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some or all variables are limited to integral or boolean (0, 1) values.
\end{define}

>>>>

\begin{define}[Node equality]
	Nodes $v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
\end{define}

\begin{define}[Value equality]
	Nodes $v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
\end{define}

@@

@@

\begin{define}[Attribute mapping]
	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $\mathcal{I} \times \mathcal{I}$ relation defined by
	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}.\]
\end{define}

@@

\begin{define}[Image of the attribute mapping]
	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$.
\end{define}

@@

\begin{define}[Candidate attribute mapping]
An attribute mapping $m$ is a~\textit{candidate attribute mapping} if it is an injective function, that is,
\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|.\]
\end{define}

@@

\begin{define}[ID set]
A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset.\]
\end{define}

@@

Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}:
\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i).\]
Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.

@@

This definition of weight for AMs or AM sets comes from \cite{fidax} again. Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.

@@

\begin{define}[Support]
\textit{Support} of an attribute mapping $m$ is defined as follows:
\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
\end{define}

@@

\begin{define}[Coverage]
\textit{Coverage} of an attribute mapping $m$ is defined as follows:
\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
\end{define}

@@

\begin{define}[Weight]
For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows:
\[weight(m) = \alpha . \phi(m) + \beta . \chi(m).\]
For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs:
\[weight(S) = \sum_{m \in S} weight(m).\]
\end{define}

@@

\begin{define}[Independent set]
Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff
\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E.\]
\end{define}

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied:
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
\end{define}

@@

It can be written in the following form:

\begin{eqnarray*}
\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
\mathbf{x} & \geqslant & 0 \\
\end{eqnarray*}
\[
\mathbf{x} =
\begin{bmatrix}
x_1 \\
x_2 \\
\ldots \\
x_n \\
\end{bmatrix},
\mathbf{c} = 
\begin{bmatrix}
c_1 \\
c_2 \\
\ldots \\
c_n \\
\end{bmatrix},
\mathbf{b} =
\begin{bmatrix}
b_1 \\
b_2 \\
\ldots \\
b_m \\
\end{bmatrix},
\]

where a minimization version is possible, too.

@@

Another way to write this formulation is this:

@@

\begin{define}[Mixed integer problem]
	MIP, or \textit{mixed integer problem}, is an~instance of~linear programming in~which some or~all variables are limited to~integral or~boolean (0, 1) values.
\end{define}

### 2011-11-19 doc/theses/vitasek/chapter-2-research.tex
Thesis: merging Lydka round 1.

According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some $K$ ($K$-\textsc{IDSet}) is in NP. Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete. The transformation from IS problem formulation to $K$-\textsc{IDSet} problem formulation is as follows.

@@

The article continues by proving that the problem of finding maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}). This again means that \textsc{Max-IDSet} is NP-complete and, furthermore, unless $\P = \NP$, \textsc{Max-IDSet} has no constant factor approximation algorithm.

@@

Fajt in \cite{fajt} summarizes several algorithms to help find XML keys in existing data, namely \textit{Gordian}, \textit{XML Primary Keys}, \textit{SPIDER} and \textit{DBA Companion}. Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in relational databases. We will describe them shortly.

@@

This is achieved by constructing a prefix tree from the tuples in the original relation, which is then pruned and traversed depth-first to find non-key attributes from which the primary keys are inferred. This algorithm still has to be adapted to search for PKs in XML data.

@@

It is possible to use a number of heuristic pruning strategies to keep the min-heap in a reasonable size. This algorithm performs very well for PKs in relational databases, however, it still has to be adapted for XML keys.

@@

Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the INDs in the database in just one pass. However, it uses a different data structure (basically a binary relation between the attributes and their corresponding values) and considers data types. Composite INDs are found using the simple ones and pruning the search space. According to the authors of \textit{SPIDER}, \textit{DBA Companion} is far inferior in performance. This algorithm has yet to be adapted to search for XML keys, too.

>>>>

According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some $K$ ($K$-\textsc{IDSet}) is in NP. Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete. The transformation from IS problem formulation to~$K$-\textsc{IDSet} problem formulation is as follows.

@@

The article continues by proving that finding the maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}). This again means that \textsc{Max-IDSet} is NP-complete and, furthermore, unless $\P = \NP$, \textsc{Max-IDSet} has no constant factor approximation algorithm.

@@

Fajt in \cite{fajt} summarizes several algorithms to help find XML keys in existing data, namely \textit{Gordian}, \textit{XML Primary Keys}, \textit{SPIDER} and \textit{DBA Companion}. Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in~relational databases. We will describe them shortly.

@@

This is achieved by constructing a prefix tree from the tuples in the original relation, which is then pruned and traversed depth-first to find non-key attributes from which the primary keys are inferred. This algorithm still has to~be~adapted to~search for PKs in XML data.

@@

It is~possible to~use a~number of~heuristic pruning strategies to~keep the~min-heap in~a~reasonable size. This algorithm performs very well for PKs in~relational databases, however, it~still has to~be~adapted for~XML keys.

@@

Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the~INDs in~the~da\-ta\-base in~just one pass. However, it~uses a~different data structure (basically a~binary relation between the~attributes and~their corresponding values) and~considers data types. Composite INDs are~found using the~simple ones and~pruning the~search space. According to~the~authors of~\textit{SPIDER}, \textit{DBA Companion} is~far inferior in~performance. This algorithm has yet to~be~adapted to~search for XML keys, too.

### 2011-11-19 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: merging Lydka round 1.

To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical'' representants.

@@

Our aim is to create a graph with approximately $v$ vertices and $e$ edges. First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}. To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others. Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}. This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.\\

@@

The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$. This is the density as the fraction of edges that are to all edges that could be in a complete graph with $|V|$ vertices.

@@

The second step is to convert the grammar into the AM model. This is done by a linear scan and retrieving a so-called \textit{flat} representation. This consists of a list of tuples in the following format.

>>>>

To interpret the data: \dataset{OVA*} sets have interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical'' representants.

@@

Our aim is to create a graph with approximately $v$ vertices and $e$ edges. First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}. To constitute an~AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others. Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of~their \texttt{attr}. This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.\\

@@

The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$. This is the density as the fraction of edges that are to all edges that could be in a~complete graph with $|V|$ vertices.

@@

The second step is to convert the grammar into the AM model. This is done by a linear scan and retrieving a so-called \textit{flat} representation. This consists of a~list of tuples in the following format.
